---
title: "Model fitting"
author: Zhang and Sheng
date: "Updated at `r lubridate::now()`"
output:
  html_notebook:
    theme: readable
    toc: true
    toc_float: true
    number_section: true
---

```{r load packages and data and basic configurations, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# laod packages
library(tidyverse)
library(lavaan)
library(semPlot)
# configurations
cfa_base <- "cfa"
data_dir <- file.path(cfa_base, "data")
mdl_spec_dir <- file.path(cfa_base, "modelspec")
res_dir <- file.path(cfa_base, "result")
# load data
ef_data <- read_csv(file.path(data_dir, "EF_Behav_scaled.csv"))
# load model specifications
mdl_three <- read_lines(file.path(mdl_spec_dir, "full_three_factor.lav"))
mdl_one <- read_lines(file.path(mdl_spec_dir, "one_factor.lav"))
mdl_shft_inhb <- read_lines(file.path(mdl_spec_dir, "ShftInhb.lav"))
mdl_shft_updt <- read_lines(file.path(mdl_spec_dir, "ShftUpdt.lav"))
mdl_updt_inhb <- read_lines(file.path(mdl_spec_dir, "UpdtInhb.lav"))
mdl_bifactor <- read_lines(file.path(mdl_spec_dir, "bifactor.lav"))
mdl_bifactor_updt_shft <- read_lines(file.path(mdl_spec_dir, "bifactor_updt_shft.lav"))
mdl_bifactor_updt_inhb <- read_lines(file.path(mdl_spec_dir, "bifactor_updt_inhb.lav"))
mdl_bifactor_inhb_shft <- read_lines(file.path(mdl_spec_dir, "bifactor_inhb_shft.lav"))
# path graph plot
cfa_plot <- function(mdl){
  semPaths(mdl, whatLabels = "par", rotation = 4, sizeMan = 10, sizeMan2 = 6.18, sizeLat = 10, mar = c(3, 20, 3, 20), nCharNodes = 0)
}
```

# Introduction

Here we try to replicate the empirical result of Miyake (2000). The data are from a project in Beijing Normal University. The following gives a first glimpse of the basic descriptive statistics.

```{r descriptive, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
psych::describe(ef_data) %>%
  select(vars, n, min, max, skew, kurtosis)
```

# Confirmatory Factor Analysis

## Three factor model

### Full three factor model

This model assumes that there exist three correlated latent variables. More specifically, the model is:

$$Inhibition =~ AntiSac + StopSignal + Stroop$$
$$Shifting =~ CateSwitch + ShiftColor + ShiftNumber$$
$$Updating =~ KeepTrack + spatialWM + WM3 $$

#### Maximum Likelihood Estimation

```{r full three factor}
fit.three.full.ml <- cfa(mdl_three, ef_data, std.lv = TRUE)
summary(fit.three.full.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.three.full.ml)
```

#### Diagnol Weighted Least Square

```{r full three factor DWLS}
fit.three.full.dwls <- cfa(mdl_three, ef_data, std.lv = TRUE, estimator = "DWLS")
summary(fit.three.full.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.three.full.dwls)
```

### Independent three factor model

This model assumes that there exist three independent latent variables.

#### Maximum Likelihood Estimation

```{r independent three factor}
fit.three.indep.ml <- cfa(mdl_three, ef_data, std.lv = TRUE, orthogonal = TRUE)
summary(fit.three.indep.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.three.indep.ml)
```

#### Diagnol Weighted Least Square

```{r independent three factor DWLS}
fit.three.indep.dwls <- cfa(mdl_three, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "DWLS")
summary(fit.three.indep.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.three.indep.dwls)
```

## One factor model

This model assumes that there exists only one latent ability.

### Maximum Likelihood Estimation

```{r one factor}
fit.one.ml <- cfa(mdl_one, ef_data, std.lv = TRUE)
summary(fit.one.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.one.ml)
```

### Diagnol Weighted Least Square

```{r one factor DWLS}
fit.one.dwls <- cfa(mdl_one, ef_data, std.lv = TRUE, estimator = "DWLS")
summary(fit.one.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.one.dwls)
```

## Two factor model

### Shifting = Inhibition

This model assumes that there exists two latent abilities. In other words, only `Updating` is the differentiated ability.

#### Maximum Likelihood Estimation

```{r two-factor shft.inhb}
fit.shft.inhb.ml <- cfa(mdl_shft_inhb, ef_data, std.lv = TRUE)
summary(fit.shft.inhb.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.shft.inhb.ml)
```

#### Diagnol Weighted Least Square

```{r two-factor shft.inhb DWLS}
fit.shft.inhb.dwls <- cfa(mdl_shft_inhb, ef_data, std.lv = TRUE, estimator = "DWLS")
summary(fit.shft.inhb.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.shft.inhb.dwls)
```

### Shifting = Updating

This model assumes that there exists two latent abilities. In other words, only `Inhibition` is the differentiated ability.

#### Maximum Likelihood Estimation

```{r two-factor shft.updt}
fit.shft.updt.ml <- cfa(mdl_shft_updt, ef_data, std.lv = TRUE)
summary(fit.shft.updt.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.shft.updt.ml)
```

#### Diagnol Weighted Least Square

```{r two-factor shft.updt DWLS}
fit.shft.updt.dwls <- cfa(mdl_shft_updt, ef_data, std.lv = TRUE, estimator = "DWLS")
summary(fit.shft.updt.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.shft.updt.dwls)
```

### Updating = Inhibition

This model assumes that there exists two latent abilities. In other words, only `Shifting` is the differentiated ability.

#### Maximum Likelihood Estimation

```{r two-factor updt.inhb}
fit.updt.inhb.ml <- cfa(mdl_updt_inhb, ef_data, std.lv = TRUE)
summary(fit.updt.inhb.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.updt.inhb.ml)
```

#### Diagnol Weighted Least Square

```{r two-factor updt.inhb DWLS}
fit.updt.inhb.dwls <- cfa(mdl_updt_inhb, ef_data, std.lv = TRUE, estimator = "DWLS")
summary(fit.updt.inhb.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.updt.inhb.dwls)
```

## Model comparison

From the above fitting results, we know that there are only two models reaching the acceptable fitting criterion. We will using **likelihood ratio test** to check which is the better model.

```{r mdoel checking}
anova(fit.three.full.dwls, fit.updt.inhb.dwls)
anova(fit.three.full.ml, fit.updt.inhb.ml)
```

From the comparisons above, we concluded that:

1. The estimation method to use is "**diagonally weighted least squares**".
1. The optimal model is **full three factor** model.

## Extract scores

We extract the scores of each subject on each ability to correlate to brain data.

```{r extract scores}
scores.dwls <- predict(fit.three.full.dwls)
write_csv(as.data.frame(scores.dwls), file.path(res_dir, "scores_dwls_bebav.csv"))
scores.ml <- predict(fit.three.full.ml)
write_csv(as.data.frame(scores.ml), file.path(res_dir, "scores_ml_behav.csv"))
```

# Bifactor Model

## Three factors

### Maximum Likelihood Estimation

```{r bifactor model ML}
fit.bifac.ml <- cfa(mdl_bifactor, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.ml)
```

### Diagnol Weighted Least Square

```{r bifactor model DWLS}
fit.bifac.dwls <- cfa(mdl_bifactor, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "DWLS")
summary(fit.bifac.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.dwls)
```

## Two factors

### Updating + Inhibition

#### Maximum Likelihood Estimation

```{r bifactor model UI ML}
fit.bifac.ui.ml <- cfa(mdl_bifactor_updt_inhb, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.ui.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.ui.ml)
```

#### Diagnol Weighted Least Square

```{r bifactor model UI DWLS}
fit.bifac.ui.dwls <- cfa(mdl_bifactor_updt_inhb, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "DWLS")
summary(fit.bifac.ui.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.ui.dwls)
```

### Updating + Shifting

#### Maximum Likelihood Estimation

```{r bifactor model US ML}
fit.bifac.us.ml <- cfa(mdl_bifactor_updt_shft, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.us.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.us.ml)
```

#### Diagnol Weighted Least Square

```{r bifactor model US DWLS}
fit.bifac.us.dwls <- cfa(mdl_bifactor_updt_shft, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "DWLS")
summary(fit.bifac.us.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.us.dwls)
```

### Shifting + Inhibition

#### Maximum Likelihood Estimation

```{r bifactor model IS ML}
fit.bifac.is.ml <- cfa(mdl_bifactor_inhb_shft, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "ML")
summary(fit.bifac.is.ml, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.is.ml)
```

#### Diagnol Weighted Least Square

```{r bifactor model IS DWLS}
fit.bifac.is.dwls <- cfa(mdl_bifactor_inhb_shft, ef_data, std.lv = TRUE, orthogonal = TRUE, estimator = "DWLS")
summary(fit.bifac.is.dwls, fit.measures = TRUE, standardized = TRUE)
cfa_plot(fit.bifac.is.dwls)
```

### Model comparison

From the above fitting results, we know that there are three models reaching the acceptable fitting criterion. We will using **likelihood ratio test** to check which is the better model.

```{r mdoel checking bifactor}
anova(fit.bifac.ml, fit.bifac.us.ml)
anova(fit.bifac.us.ml, fit.bifac.is.ml)

anova(fit.bifac.us.dwls, fit.bifac.is.dwls)
```
## Extract scores

We extract the scores of each subject on each ability to correlate to brain data.

```{r extract bifactor scores}
scores.dwls <- predict(fit.bifac.dwls)
write_csv(as.data.frame(scores.dwls), file.path(res_dir, "scores_dwls_bifactor_behav.csv"))
scores.ml <- predict(fit.bifac.ml)
write_csv(as.data.frame(scores.ml), file.path(res_dir, "scores_ml_bifactor_behav.csv"))
```
